Lattice Boltzmann simulations of magnetically responsive\\ ellipsoidal particle stabilized emulsions

### 2025-03-23 chapters/dissertation-tap.tex
Changes 1. Working on feedback     a. Changed abstrat     b. Cleaned up bibliography for diagnosis if issues were there     c. Fixed reproduction licenses in figure captions

\prefacesectiontoc{Abstract}
%\input{abstract} % This file (abstract.tex) contains the text
                 % for an abstract or type text here.

Porous materials have found great importance in applications such as water filtration, catalyst supports, battery 
electrodes and bio engineered materials. To accommodate a greater variety of pore sizes and continuous fabrication 
techniques with minimal waste, bottom up synthesis techniques have been gaining in popularity. One such technique is
emulsion templating of a specific microstructure known as a bicontinuous interfacially jammed emulsion gels(bijels).

However, casting mixture independent microstructure control is an open issue, especially applicable in continuous 
bijel synthesis strategies such as Solvent Transfer Induced Phase separation and as in-situ control mechanisms to
control the microstructure of a bijel to control properties such as separation efficiency. Stimuli response has been identified as a 
technique to enable this. This study focuses on utilizing stimuli response of bijels stabilized with ellipsoidal microparticles, 
and the resulting microstructure and rheological implications this induces.

We first identify that upon application of a field to a bijel stabilized with ellipsoidal particles during synthesis, 
microstructure change is observed, not present when analyzing bijels stabilized with spherical particles. Upon 
application of the field, particles reorient to the direction of the field, causing directional jamming that creates
microstructure anisotropy.

We then identify that bijels stabilized with ellipsoidal particles have tunable microstructures after jamming. 
This process is dependent upon the initial particle order present in the bijel, as well as the difference 
between the initial and final particle order. We also show that the structural response observed is state 
dependent, meaning that the time of magnetic field application affects the microstructure change observed.

Finally, we characterize that bijels retain their shear thinning behavior when analyzing the rheological response of bijels
under magnetic fields. We identify that the particle ordering plays a role in the evolution of the microstructure under shear
and the rheological response.

\end{front}

>>>>

\prefacesectiontoc{Abstract}
%\input{abstract} % This file (abstract.tex) contains the text
                 % for an abstract or type text here.
    Porous materials are important in a wide range of applications, including water filtration, 
    catalyst supports, battery electrodes, and bioengineered materials. To accommodate a broader range of pore 
    sizes and enable scalable fabrication with minimal waste, bottom-up synthesis techniques have gained 
    increasing attention. Among these, emulsion templating based on bicontinuous interfacially jammed 
    emulsion gels (bijels)—has emerged as a promising approach due to its ability to produce complex, 
    interconnected porous structures.

    Despite these advantages, achieving microstructural control independent of the casting mixture remains a 
    key challenge, particularly in continuous bijel fabrication methods such as Solvent Transfer Induced Phase 
    Separation (STrIPS). Additionally, in situ control strategies for tailoring properties such 
    as separation efficiency can be useful additions to control material performance. 
    Stimuli-responsive approaches have been identified as a potential route to 
    overcome these constraints. This study focuses on the magnetic stimuli-response of bijels stabilized by 
    magnetically responsive ellipsoidal microparticles, examining both the resulting microstructural 
    modifications and their rheological implications.

    It is shown that the application of a magnetic field during bijel formation induces microstructural 
    changes in systems stabilized by ellipsoidal particles—effects not observed in bijels stabilized by 
    spherical particles. The applied field causes particle reorientation along its direction, leading to 
    directional jamming and the development of anisotropic microstructures.

    Post-synthesis, bijels stabilized by ellipsoidal particles exhibit tunable microstructure. This responsiveness 
    is governed by the degree of particle order present prior to field application, as well as the extent of 
    reordering induced by the field. The system exhibits state-dependent behavior, indicating that the timing of 
    magnetic field application significantly influences the magnitude of the microstructural response characterized.

    Rheological characterization reveals that bijels retain their shear-thinning behavior under applied magnetic 
    fields. However, the degree of particle ordering is found to influence both the evolution of the microstructure 
    under shear and the corresponding rheological response, highlighting the interplay between field-induced 
    structuring and mechanical performance.
\end{front}

### 2025-03-23 chapters/introduction.tex
Changes 1. Addressed some image sizes and phrasing consistency
Draft 1 complete
Changes 1. Working on feedback     a. Changed abstrat     b. Cleaned up bibliography for diagnosis if issues were there     c. Fixed reproduction licenses in figure captions

\begin{figure}
    \centering
    \includegraphics[scale = 0.5]{figures/introduction/bijel_templating.png}
    \caption{Fabrication of a graphite oxide battery electrode using bijel template. Reproduced from Garcia et al. 
             Nanoscale Adv., 2019,1, 3870-3882 under license number 1548092-1. \cite{garcia_scalable_2019}}
    \label{fig:bijel_template}
\end{figure}

@@

\begin{figure}[h]
    \centering
    \includegraphics[scale = 4]{figures/introduction/STrIPS.jpg}
    \caption{STrIPS in action. Extrusion of the bijel casting mixture into a non-solvent bath, followed by removal of solvent from 
             the casting mixture through diffusion. \cite{haase_continuous_2015} Reproduced from Haase et al. Adv. Mater., 
             Volume27, Issue44, 2015, Pages 7065-7071 under license number 5913140219015. }
    \label{fig:strips}
\end{figure}

@@

\begin{figure}[h]
    \centering
    \includegraphics[scale = 1]{figures/introduction/magnetophoresis_emulsion.jpeg}
    \caption{Enhanced oil recovery of oil from water using magnetically responsive emulsion stabilizers, allowing for 
             locomotion and controlled coalescence of cargo from a matrix phase using magnetic fields. \cite{tham_magnetophoresis_2021} 
             Reproduced from Tham et al. Langmuir 2021, 37, 5, 1811-1822 under the ACS standard license.}
    \label{fig:magnetophoresis_droplet}
\end{figure}

@@

Building on this background, we propose the use of anisotropic particles as a technique to enable microstructure modification of bijels upon the application 
of an external field, both during and after fabrication. This work aims to explore methods for controlling the bijel microstructure independently of the casting 
mixture composition by utilizing magnetic stimuli on bijels stabilized with magnetically responsive ellipsoidal particles.  
The proposed mechanism of microstructure modification is based on the reorientation of particles in response to the applied magnetic field, which alters the 
packing arrangement of the particles at the interface. By modifying the interfacial packing, the point at which the particles jam is also altered, thereby 
affecting the jamming point of the bijel. This mechanism is expected to be effective both during and after the synthesis process. Additionally, the 
processability of bijels produced using this technique will be examined to assess how changes in particle packing influence the rheological properties of 
the material.

\section{Research objective}

% this work seeks to identify means to control the bijel microstructure independent of the casting mixture composition utilizing magnetic stimuli response of anisotropic particles. The proposed mechanism of microstructure modification is driven through reorientation of the particles to the applied magnetic field altering how the particles pack on the interface.

\subsection{Aim 1: Determination of the degree of microstructure changes expected when applying a constant field during bijel formation}
\label{section:aim1_desc}

The jamming point of a bijel plays a crucial role in determining the resulting microstructure, as demonstrated by previous studies comparing the length scales 
of bijels synthesized with varying particle volume fractions and sizes \cite{jansen_bijels_2011, reeves_particle-size_2015}. Additionally, it has been shown 
that magnetic fields can facilitate the self-assembly of anisotropic particles at interfaces and control the interfacial angle between the particle's axis 
of symmetry and the interface \cite{davies_interface_2014, davies_assembling_2014}. However, on a curved interface with numerous other particles, it remains 
unclear how the orientation of multiple particles in response to the field and their regular arrangement will influence the jamming behavior of the bijel 
\cite{bresme_orientational_2007, davies_interface_2014}.

To evaluate the microstructure obtained upon the application of a field during the formation of the bijel, homogeneous mixtures containing spherical and 
ellipsoidal particles will be subjected to field strengths both above and below the critical field strength. This critical field strength is the threshold 
at which particle orientations to the interface are governed by the applied field, as calculated by Bresme and Faraudo for the particle aspect ratio 
\cite{bresme_orientational_2007, davies_interface_2014}. The range of selected field strengths will provide insight into the mechanisms controlling 
microstructural changes, enabling verification of the hypothesis and identification of the conditions under which the most control over the microstructure 
can be achieved.

>>>>

\begin{figure}
    \centering
    \includegraphics[scale = 0.5]{figures/introduction/bijel_templating.png}
    \caption{Fabrication of a graphite oxide battery electrode using bijel template. Used with permission of 
    Royal Society of Chemistry, from Scalable synthesis of gyroid-inspired freestanding three-dimensional graphene architectures, 
    Garcia et al., 1, 10, 2018 permission conveyed through Copyright Clearance Center, Inc}
    \label{fig:bijel_template}
\end{figure}

@@

\begin{figure}[h]
    \centering
    \includegraphics[scale = 5]{figures/introduction/STrIPS.jpg}
    \caption{STrIPS in action. Extrusion of the bijel casting mixture into a non-solvent bath, followed by removal of solvent from 
             the casting mixture through diffusion. \cite{haase_continuous_2015}. Reproduced with permission from Wiley-VCH Verlag GmbH 
             \& Co. KGaA from \textit{Continuous Fabrication of Hierarchical and Asymmetric Bijel Microparticles, Fibers, and Membranes by 
             Solvent Transfer-Induced Phase Separation (STRIPS)}, Haase, M. F., Stebe, K. J., 
             \& Lee, D., \textit{Advanced Materials}, 27(44), 7065-7071, 2015}
    \label{fig:strips}
\end{figure}

@@

\begin{figure}[h]
    \centering
    \includegraphics[scale = 1.5]{figures/introduction/magnetophoresis_emulsion.jpeg}
    \caption{Enhanced oil recovery of oil from water using magnetically responsive emulsion stabilizers, allowing for 
             locomotion and controlled coalescence of cargo from a matrix phase using magnetic fields. \cite{tham_magnetophoresis_2021} 
             K. Tham, W. M. Ng, S. S. Leong, S. P. Yeap, S. C. Low, H. L. Lee and J. Lim. Magnetophoresis of magnetic pickering 
             emulsions under low field gradient: Macroscopic and microscopic motion. 37, 1811: Reprinted (adapted) with permission from 
             Langmuir 2021, 37, 5, 1811-1822. Copyright 2025 American Chemical Society.}
    \label{fig:magnetophoresis_droplet}
\end{figure}

@@

Building on this background, the use of anisotropic particles, specifically ellipsoidal particles, is proposed as a technique to enable microstructure modification 
of bijels upon the application 
of an external field, both during and after fabrication. This work aims to explore methods for controlling the bijel microstructure independently of the casting 
mixture composition by utilizing magnetic stimuli on bijels stabilized with magnetically responsive ellipsoidal particles.  
The proposed mechanism of microstructure modification is based on the reorientation of particles in response to the applied magnetic field, which alters the 
packing arrangement of the particles at the interface. By modifying the interfacial packing, the point at which the particles jam is also altered, thereby 
affecting the jamming point of the bijel. This mechanism is expected to be effective both during and after the synthesis process. Additionally, the 
processability of bijels produced using this technique will be examined to assess how changes in particle packing influence the rheological properties of 
the material.

% \section{Research objective}

% % this work seeks to identify means to control the bijel microstructure independent of the casting mixture composition utilizing magnetic stimuli response of anisotropic particles. The proposed mechanism of microstructure modification is driven through reorientation of the particles to the applied magnetic field altering how the particles pack on the interface.

% \subsection{Aim 1: Determination of the degree of microstructure changes expected when applying a constant field during bijel formation}
% \label{section:aim1_desc}

The jamming point of a bijel plays a crucial role in determining the resulting microstructure, as demonstrated by previous studies comparing the length scales 
of bijels synthesized with varying particle volume fractions and sizes \cite{jansen_bijels_2011, reeves_particle-size_2015}. Additionally, it has been shown 
that magnetic fields can facilitate the self-assembly of ellipsoidal particles at interfaces and control the interfacial angle between the particle's axis 
of symmetry and the interface \cite{davies_interface_2014, davies_assembling_2014}. However, on a curved interface with numerous other particles, it remains 
unclear how the orientation of multiple particles in response to the field and their regular arrangement will influence the jamming behavior of the bijel 
\cite{bresme_orientational_2007, davies_interface_2014}. Therefore the microstructural response of bijels as a function of the applied field will be 
characterized.

A homogeneous mixtures containing spherical and ellipsoidal particles will be subjected to field strengths both above and below the critical field strength. 
This critical field strength is the threshold 
at which particle orientations to the interface are governed by the applied field, as calculated by Bresme and Faraudo for the particle aspect ratio 
\cite{bresme_orientational_2007, davies_interface_2014}. The range of selected field strengths will provide insight into the mechanisms controlling 
microstructural changes, enabling verification of the hypothesis and identification of the conditions under which the most control over the microstructure 
can be achieved.

### 2025-03-23 chapters/lit_review.tex
Changes 1. Working on feedback     a. Changed abstrat     b. Cleaned up bibliography for diagnosis if issues were there     c. Fixed reproduction licenses in figure captions

\begin{figure}
    \centering
    \includegraphics[scale = 0.5]{figures/literature_review/interfacial_curvature.png}
    \caption{Quadropolar capillary interactions around prolate ellipsoidal particles caused by interfacial deformations 
             around the particle. \cite{loudet_capillary_2005} Reproduced from Loudet et al. Phys. Rev. Lett. 2005, 94, 018301
             with license number RNP/25/FEB/088185}
    \label{fig:anisotropic_particle_interface}
\end{figure}

@@

\begin{figure}
    \centering
    \includegraphics[scale = 0.3]{figures/literature_review/state_diagram.jpg}
    \caption{Formation criteria for a particle stabilized emulsion that can form bijel. \cite{velankar_non-equilibrium_2015}
            Reproduced from Velankar Soft Matter, 2015,11, 8393-8403 using license number 1551968-1.}
    \label{fig:state_diagram_particle_emulsions}
\end{figure}

@@

\begin{figure}
    \centering
    \includegraphics[scale = 3]{figures/literature_review/bijel_under_shear.jpeg}
    \caption{Schematic of a bijel with hard-sphere particles undergoing shear, demonstrating migration and detachment of 
             particles at the interface. \cite{bonaccorso_shear_2020} Reproduced from Bonaccorso et al. AIP Advances 1 
             September 2020; 10 (9): 095304, under the Creative Commons CC BY License.}
    \label{fig:bijel_under_shear}
\end{figure}

>>>>

\begin{figure}
    \centering
    \includegraphics[scale = 0.5]{figures/literature_review/interfacial_curvature.png}
    \caption{Quadropolar capillary interactions around prolate ellipsoidal particles caused by interfacial deformations 
             around the particle. \cite{loudet_capillary_2005} Reprinted Figure 4 with permission from
             Loudet et al, Phys. Rev. Letter., 94, 08301, 2003 Copyright 2025 by the American Physical Society.}
    \label{fig:anisotropic_particle_interface}
\end{figure}

@@

\begin{figure}
    \centering
    \includegraphics[scale = 0.3]{figures/literature_review/state_diagram.jpg}
    \caption{Formation criteria for a particle stabilized emulsion that can form bijel. \cite{velankar_non-equilibrium_2015}
            Used with permission of Royal Society of Chemistry, from A non-equilibrium state diagram for liquid/fluid/particle 
            mixtures, Velankar, 1, 11, 2015. Permission conveyed through Copyright Clearance Center, Inc}
    \label{fig:state_diagram_particle_emulsions}
\end{figure}

@@

\begin{figure}
    \centering
    \includegraphics[scale = 3]{figures/literature_review/bijel_under_shear.jpeg}
    \caption{Schematic of a bijel with hard-sphere particles undergoing shear, demonstrating migration and detachment of 
             particles at the interface. \cite{bonaccorso_shear_2020} Reproduced from Shear dynamics of confined bijels
             ,Bonaccorso et al. AIP Advances 1, September 2020; 10 (9): 095304, under the Creative Commons CC BY License.}
    \label{fig:bijel_under_shear}
\end{figure}
